---
title: "automatic_celltype_methods"
author: "Marcos Machado"
date: "12/24/2021"
output: html_document
---

### Not working cell type identification methods 
#### Cell type identification using scCATCH
(this is not working properly. There are a lot of bugs in the CellMatch database that do not let me simultaneously use the cancer option and also identify healthy tissue)


Tissue and cancer types acording to the CellMatch database. Make sure to only include up-regulated genes
```{r}
# include only positively regulated gene set (very important)
markersAll_upregulated <- FindAllMarkers(mergedPCA, assay = "SCT", return.thresh = 0.01, only.pos = TRUE)
# markers_scCATCH <- findmarkergenes(mergedPCA, species = 'Human', cluster = 'All', match_CellMatch = TRUE, tissue = c("Breast", "Mammary epithelium", "White adipose tissue"), pvalue = 0.05)

clu_ann <- scCATCH(object = markersAll_upregulated,
                   species = 'Human',
                   tissue = c("Breast")
                   # cancer = c("Breast Cancer")
                   )
```


#### Cell type identification using CellAssign
(not working because of problems loadng Tensorflow :( )

This is a cluster-independent method to assign cell types based on the gene expression matrix

##### Data preparation
Reading CellMarker database file with all cell types and marker genes, and generating a gene-by-celltype binary matrix. Also, we perform a conversion of our Seurat object to a SingleCellExperiment (SCE) object, which is now compatible with CellAssign
```{r}
# loading database file to data.frame
CellMarkersDF <- read.csv("/home/marcos/Downloads/Single_cell_markers.txt", sep = "\t")
CellMarkersDF <- CellMarkersDF[CellMarkersDF$speciesType == "Human",] # Remove mouse rows
CellMarkersDF$cellName <- paste(CellMarkersDF$cellName, "_", CellMarkersDF$tissueType, "_", CellMarkersDF$cellType, "_", CellMarkersDF$PMID)

# extracting celltypes and markers to binary matrix

CellMarkers <- list(CellMarkersDF$cellName)
CellMarkers[CellMarkersDF$cellName] = strsplit(as.character(CellMarkersDF$cellMarker), ", ")
CellMarkers <- CellMarkers[-1]
CellMarkers <- marker_list_to_mat(CellMarkers)


# SCE conversion
mergedPCA.sce <- as.SingleCellExperiment(mergedPCA, assay = "RNA")

# mergedPCA.sce <- SingleCellExperiment(assays = list(counts = GetAssayData(object = mergedPCA, slot = "data")))
```

##### Run CellAssign

```{r}
# Intersection between gene-by-cellType matrix and sce genes
shared <- intersect(rownames(CellMarkers), rownames(mergedPCA.sce))
mergedPCA.sce <- computePooledFactors(mergedPCA.sce, assay.type = "counts")
s <- sizeFactors(mergedPCA.sce)

fit <- cellassign(exprs_obj = mergedPCA.sce[shared, ], 
                  marker_gene_info = CellMarkers[shared, ], 
                  learning_rate = 1e-2, 
                  s = s,
                  shrinkage = TRUE,
                  verbose = FALSE)
```
